Retrofitting Effect Handlers onto OCaml

### 2020-11-17 retro-concurrency.tex
edits

Kawahara et al.~\cite{Kawahara20} implement one-shot effect handlers using
coroutines as a macro-expressible translation, and present an embedding in Lua
and Ruby. Lua provides asymmetric coroutines~\cite{Lua} where each coroutine
uses its own stack similar to how each handled computation runs in its own
fiber in Multicore OCaml. Lua does not allow coroutines to include C frames, as in
Multicore OCaml. 
%It is useful to note that Lua is an interpreted
%language, unlike OCaml, and DWARF unwind tables are not a concern in Lua.

@@

Figure~\ref{res:macro} presents the performance of macro benchmarks that do not
use effects. The aim is to quantify the impact of the addition of effect
handlers to the language on code that does not use effect handlers. The |MC| is
Multicore OCaml with default red zone size (16 words). |MC+RedZone0| uses red
zone size of 0; every OCaml function has a stack overflow check. |MC+RedZone32|
uses a red zone size of 32 words.

@@

Languages with effect handlers typically allow continuations to be resumed more
than once to enable use cases such as backtracking search and probabilistic
programming (Uber's Pyro langauge~\cite{Pyro}). Multicore OCaml has an
experimental feature for explicitly cloning continuations, which makes a copy
of each of its constituent fibers. Multi-shot continuations are difficult to
reason about in the presence of linear resources such as file descriptors.
Recall that OCaml programs are written with the expectation that function calls
return \emph{exactly-once} either normally or exceptionally. With multi-shot
continuations, they may return \emph{more than once}, leading to errors such as
double-freeing malloced memory. Since our goal is to retrofit effect handlers
onto OCaml, we relegate continuation cloning to the OCaml |Obj| module, which
contains unsafe features.

>>>>

Kawahara et al.~\cite{Kawahara20} implement one-shot effect handlers using
coroutines as a macro-expressible translation, and present an embedding in Lua
and Ruby. Lua provides asymmetric coroutines~\cite{Lua} where each coroutine
uses its own stack similar to how each handled computation runs in its own
fiber in Multicore OCaml.
%Lua does not allow coroutines to include C frames, as in
%Multicore OCaml.  - avsm/kc not quite, see https://coco.luajit.org/index.html
%It is useful to note that Lua is an interpreted
%language, unlike OCaml, and DWARF unwind tables are not a concern in Lua.

@@

Figure~\ref{res:macro} presents the performance of macro benchmarks that do not
use effects. The aim is to quantify the impact of the addition of effect
handlers to the language on code that does not use effect handlers. The |MC| is
Multicore OCaml with the default red zone size (16 words). |MC+RedZone0| uses red
zone size of 0; every OCaml function has a stack overflow check. |MC+RedZone32|
uses a red zone size of 32 words.

@@

Languages with effect handlers typically allow continuations to be resumed more
than once to enable use cases such as backtracking search and probabilistic
programming (Uber's Pyro language~\cite{Pyro}). Multicore OCaml has an
experimental feature for explicitly cloning continuations, which makes a copy
of each of its constituent fibers.

Multi-shot continuations are difficult to
reason about in the presence of linear resources such as file descriptors.
Recall that OCaml programs are written with the expectation that function calls
return \emph{exactly-once} either normally or exceptionally. With multi-shot
continuations, they may return \emph{more than once}, leading to errors such as
double-freeing malloced memory. Since our goal is to retrofit effect handlers
onto OCaml, we relegate continuation cloning to the OCaml |Obj| module, which
contains unsafe features.
